% Introduce a chapter describing LSTM and GRU, publications to review and accurate comparison.
%
Following Literature Review, this chapter will analyse the selected branch of Neural Networks in detail.
It explores the Recurrent NN in-depth and establishes the methodology for further research and implementation
The accuracy against complexity plot, Figure~\ref{fig:alg-comparioson}, states that there are two types of RNN worth detailed exploration and implementation: Long Short-Term Memory (LSTM) and Gated Recurrent Unit (GRU).
Each type will be the subject of a detailed review with multiple published papers analysed, implemented and compared with percentage accuracy. 

% History of using some of those types of models
%
Machine learning methods can establish relationships in complicated, multi-dimensional non-linear systems~\cite{hansen_support_2005,anton_battery_2013,he_state_2014}.
Many attempts to implement different NN exist, but the most promising were explored by following papers using Recurrent Neural Networks~\cite{song_lithium-ion_2018,Chemali2017,mamo_long_2020,jiao_gru-rnn_2020,xiao_accurate_2019,javid_adaptive_2020,zhang_deep_2020}, invlolving modifying models' structure, optimising fitting process and enhancing data pipelines.
RNN effectiveness in time-series dependant problems was shown using internal neurons to process data sequences with varying lengths~\cite{}\textcolor{red}{reputable machine learning/neural networks venue (NIPS, ICML, etc.)}.
The cells act as memory units, building relationships and giving output based on multiple inputs over time.

% What is terrible about RNN models, and what did other authors improve them?
%
Over the past years, the RNN approach has received extensive attention from different researchers. The most recent attempt to determine the Li-Ion battery's remaining useful life was the Gated Recurrent Unit models~\cite{song_lithium-ion_2018,javid_adaptive_2020,xiao_accurate_2019,jiao_gru-rnn_2020}.
The earliest approach utilised a battery charge's regression nature only using stateless models~\cite{song_lithium-ion_2018,jiao_gru-rnn_2020,xiao_accurate_2019} when input impact is not preserved for further predictions.
Later, some approaches introduced additional parameters to support the NN learning process~\cite{mamo_long_2020,jiao_gru-rnn_2020,javid_adaptive_2020}.
Besides good converges within 2-5\% to the actual data, the model is used to determine critical events, like the time before complete charge depletion or overcharge.
However, it did not receive wide application due to the battery's initial state requirement and prior utilisation as input features.
The most popular way lies in determining the charge's value using a fixed size history of voltage, current and temperature in stateless Long Short-Term Memory (LSTM) models~\cite{Chemali2017,mamo_long_2020,javid_adaptive_2020,zhang_deep_2020}.
The method's advantage is independent of the charge or discharge cycles at different times, as long as the history samples are in order.
However, estimation determines values based on fixed history samples and does not preserve for the next prediction, making practical estimation of single or several charge values, but not determining critical events ahead.
That can be achieved with stateful models, which preserve and update the state within LSTM cells themself with every passing prediction, making the impact of the history affect results continuously.

% Validation procedure problem and lack of data of the same condition
%
Both stateful and stateless methods rely on input samples' quality and length.
Chemali et al.~\cite{Chemali2017} researched the impact of the history length of input samples: the longer the period of input readings, the better accuracy model produced and the longer it takes to compute.
\textcolor{red}{Worth bringing his data in figure form in this part from article part}
The result of the research is the root square parabola behaviour of size of the history against error in the prediction.
It was concluded that the length of five minutes made the most optimal trade-off between computational weight and general accuracy.
In addition, the normalisation technique is commonly used to speed up the fitting process, such as Standardisation, Min-Maxing or Scaling data between two constants.
In the SoC estimation, scaling temperatures closer to the voltage and current reduce the time before the first valid results.
After identifying the most optimal input preferences without significant effect on the performance, one of the further developments applies changes to the structure of GRU or LSTM by adding additional layers (Attention or Extra Dense layers)~\cite{mamo_long_2020, jiao_gru-rnn_2020}.
Another way is to change the optimisation process to achieve similar accuracy faster (i. e., adding a Momentum algorithm to the Stochastic Gradient Optimisations process or replacing gradient calculations with statistical)~\cite{xiao_accurate_2019, javid_adaptive_2020}.
Those methods modify standard ways introduced earlier in model training by applying additional operations.
As a result, they met to achieve better accuracy faster using the same or similar training approaches. 

% Online validation attempt and what good about it
%
The drawback lies within the validation procedures of produced models.
Ideally, this process must be performed with similar data but under different conditions or at different times.
However, once the produced model is taken through different unseen scenarios, the accuracy lowers drastically compared to early reported results since it starts to experience something completely unexpected.
The lack of training samples or means to produce under different conditions affected the final judgment appears in many publications. 
Even though the estimation produces accurate output with tabled data from precisive battery cycling machines placed under actual driving conditions - the model will have difficulties matching training performance. 

%
%
% Several attempts to introduce an online procedure for model performance measurement have been integrated into the training process to solve this problem.
% By not being limited to the battery testing machine's table data, i.e. the validation mechanism to tune an NN model based on battery data during actual battery cycling, researchers attempted to generalise the fitting process as best as hardware allowed~\cite{zhang_deep_2020}.
% This method brings the model learning process closer to real-time battery utilisation without adding modelling complexity.

% Tabled summary from reviewed articles
%
Table~\ref{tab:review} summarises the most common methods applied to SoC estimation, highlighting the model cell type, structure to define input sample type, optimiser selection, and additional features introduced by authors to improve predictions.
The model type highlights a primary path in structuring a Neural Net model.
% Statefulness defines the input method, where stateless uses a fixed size of input samples per feature and statefully applies each time sample at a time for all features one by one.
Optimiser selection sets the algorithm for the learning process from one of the following methods: Adaptive moment estimation (Adam), Nesterov adaptive moment estimation  (Nadam), Stochastic gradient descent (SGD), AdaMax (AM) and Differential Evolution (DE).
\begin{table*}[h]
    \caption{Reviewed ML implementations for SoC estimation.}
    \label{tab:review}
\resizebox{\textwidth}{!}{
\begin{tabular}{p{2.0cm}|p{0.8cm}|p{1.0cm}|p{0.8cm}|p{0.8cm}|p{1.0cm}|p{1.0cm}|p{0.9cm}|p{0.7cm}|p{0.7cm}|p{2.1cm}}
    %\multicolumn{12}{c}{Unknown yet table} \\
    \hline
    \multicolumn{1}{ c }{} & 
    \multicolumn{2}{|c|}{Model} & 
    \multicolumn{2}{ c|}{State-} & 
    \multicolumn{5}{ c|}{Optimiser} &
    \multirow{2}{ 4em }{Extension} \\
    \cline{1-4} \cline{5-10}
    Ref & GRU  & LSTM & -less & -ful & Adam & Nadam & SGD & AM & DE &           \\
    \hline
    Song~\cite{song_lithium-ion_2018}
        & \chk &      &       & \chk  & \chk &       &     &    &    & 4 Layers  \\
    Chemali~\cite{Chemali2017}
        &      & \chk & \chk  &       & \chk &       &     &    &    &           \\
    Mamo~\cite{mamo_long_2020}
        &      & \chk &  \chk &       &      &       &     &    &\chk& Attention \\
    Jiao~\cite{jiao_gru-rnn_2020}
        & \chk &      &       & \chk  &      &       & \chk&    &    & Momentum  \\
    Xiao~\cite{xiao_accurate_2019}
        & \chk &      &       & \chk  &      & \chk  &     &\chk&    & Ensemble  \\
    Javid~\cite{javid_adaptive_2020}
        & \chk &      & \chk  &       & \chk &       &     &     &   & Robust    \\
    Zhang~\cite{zhang_deep_2020}
        &      & \chk & \chk  &       &      & \chk  &     &     &   & Online    \\
    \hline
\end{tabular}
}
\end{table*}

% What has been done so far by other authors, and what are they using to achieve that?
%
All reviewed authors, \mbox{Table~\ref{tab:review}} conducted model experiments on battery cycling data of different cell types.
Most used table data of real-time sensory results from battery cyclers to validate efficiencies.
Data gets generated using different current schedulers, also called driving profiles.
An equally time-based sample of current consumption acts as an input to the battery cycler, intended to recreate a stress test on a battery or human driving behaviour.
There are three the most commonly used in the research of this area: Dynamic Stress Test (DST), Aggressive Driving (US06) and Federal-Urban Driving Scheduler (FUDS).
Unlike some general simple static discharge processes, which commonly appear in other battery-based tools, driving profiles include some amount of regenerative driving to simulate the actual application of the battery for an electric vehicle.

% How the performance testing is performed and what challenges it brings.
%
Mamo et al.~\cite{mamo_long_2020} conducted experiments by validating the RNN model performance by training one driving profile and testing against another.
The results showed an increase in Root Mean Squared Error compared to the single dataset training and validation scenario.
There have not been many similar experiments against other neural nets or SoC estimation methods.
A single profile's battery cycling data testing may not validate ML methods' efficiency in driving an electric vehicle.
The inability to determine the charge's current state in EV applications makes the online learning process inapplicable.
Even with other SoC estimation technique usage, the computational complexity of training any NN is loaded to be performed on an accumulator's battery management system.
An offline trained model had the advantage of insignificant resource consumption during the prediction stage, making it a prefered way for an EV.
All further model testing will be applied through varying battery cycling profiles to capture the influence of data type to model efficiency.

% The goal of this chapter. It will investigate, implement and compare.
%
This chapter investigates, implements and compares memory-based models of RNN to predict the State of Charge and additional built-on over-time techniques to select the most effective and least resource requiring onboard-based computations appliable within Electrical Vehicles.
% It would attempt to identify potential flaws in preparing battery models and introduce potential solutions to improve upon.
The recent advances and commonly used subsets are Gated Recurrent Unit and Long Shot-Term Memory unit cells.
Recurrent Neural Network has been confirmed to be suitable for the battery-related system by authors, such as Chemali ~\cite{LSTM_Hochreiter1997}.
However, there is no valid proof of which GRU or LSTM is better for a battery SoC estimation and what would be the best set of hyper-parameters to train a model.
The least time-consuming way would be to set up a methodology by trying out one of the models with different parameter combinations and calculating average accuracy across multiple attempts.
The selected set for the EV scenario will be applied to every investigated model and compared evenly between each other.
Each subset will contain implementation from various articles changing either structure of the models or learning approaches.
Ultimately, each method will be taken through the same battery data of DST, US06 and FUDS driving profiles and compared against the robustness and accuracy of the estimation State of Charge of Lithium-Ion batteries.

%
%
% The remaining sections are organised as follows: all methodology, methods and data discussed in Section~\ref{sec:Body}.
% The details of data management and model structure are in Subsection~\ref{subsec:RNN}.
% Subsections~\ref{subsec:structure},~\ref{subsec:optimisers} and~\ref{subsec:soft} separate all details for each GRU and LSTM method using multiple optimising algorithms.
% Section~\ref{sec:conclussion} gives the results of implementation, performance, and discussion and concludes the critical analysis.